Scalable Extreme Deconvolution

### 2019-10-20 ml4ps_paper/paper.tex
Change to camera-ready and edits per reviewer comments.

% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage{neurips_2019}

@@

The aim of extreme deconvolution is to perform density estimation on a noisy $d$-dimensional dataset $\{\bx_i\}_{i=0}^N$, where $\bx_i$ was generated by adding zero-mean Gaussian noise $\epsilon_i$ with known per-datapoint covariance $S_i$ to a projection $R_i$ of a true value $\bv_i$,
\begin{equation}
  \bx_i = R_i\bv_i + \epsilon_i,\quad  \epsilon_i \sim \mathcal{N}(\mathbf{0}, S_i).
\end{equation}
We assume that $\bv_i$ can be modelled by a mixture of Gaussians with $K$ components,
\begin{equation}
p(\bv_i \mid \theta) = \sum_j^K \alpha_j \,\mathcal{N}(\bv \mid \bm_j, V_j), \quad \theta = \{\alpha_j, \bm_j, V_j\}_{j=1} ^ K.
\end{equation}
As the noise model is Gaussian and the model of the underlying density is a mixture of Gaussians, the probability of $\bx_i$ is also a Gaussian mixture.
The total log-likehood of the model is
\begin{equation}
\mathcal{L}(\theta) = \sum_i^N \log \sum_j^K \alpha_j\,\mathcal{N}(\bx_i \mid R_i\bm_j, T_{ij}), \quad T_{ij} = R_iV_jR_i^T + S_i.
\end{equation}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/learning.pdf}
  \caption{\textbf{Left}: Average training log-likelihood as a function of training epoch for models with $K=256$. Epochs rescaled by average training time. Error bars not visible. \textbf{Right}: Training time as a function of mixture components $K$. Error bars indicate $\pm$ 2 standard deviations.}
  \label{fig:training}
\end{figure}

@@

>>>>

% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage[final]{neurips_2019_ml4ps}

@@

The aim of extreme deconvolution is to perform density estimation on a noisy $d$-dimensional dataset $\{\bx_i\}_{i=0}^N$, where $\bx_i$ was generated by adding zero-mean Gaussian noise $\epsilon_i$ with known per-datapoint covariance $S_i$ to a projection $R_i$ of a true value $\bv_i$,
\begin{equation}
  \bx_i = R_i\bv_i + \epsilon_i,\quad  \epsilon_i \sim \mathcal{N}(\mathbf{0}, S_i).
\end{equation}
We assume that $\bv_i$ can be modelled by a mixture of Gaussians with $K$ components,
\begin{equation}
p(\bv_i \mid \theta) = \sum_j^K \alpha_j \,\mathcal{N}(\bv \mid \bm_j, V_j), \quad \theta = \{\alpha_j, \bm_j, V_j\}_{j=1} ^ K,
\end{equation}
parameterised by mixture weight $\alpha_j$, mean $\bm_j$ and covariance $V_j$.
As the noise model is Gaussian and the model of the underlying density is a mixture of Gaussians, the probability of $\bx_i$ is also a Gaussian mixture.
The total log-likehood of the model is
\begin{equation}
\mathcal{L}(\theta) = \sum_i^N \log \sum_j^K \alpha_j\,\mathcal{N}(\bx_i \mid R_i\bm_j, T_{ij}), \quad T_{ij} = R_iV_jR_i^T + S_i.
\end{equation}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/learning.pdf}
  \caption{\textbf{Left}: Average training log-likelihood as a function of training on the Gaia subset for models with $K=256$. Epochs rescaled by average training time. Error bars not visible. \textbf{Right}: Training time as a function of mixture components $K$. Error bars indicate $\pm$ 2 standard deviations.}
  \label{fig:training}
\end{figure}

@@

\subsubsection*{Acknowledgments}
We are grateful to David W. Hogg for providing us with the problem and advising us on our experimental setup.
Our experiments made use of AstroPy~\citep{astropy:2013,astropy:2018}, corner.py~\citep{corner}, Numpy~\citep{numpy}, Pandas~\citep{mckinney-proc-scipy-2010}, PyTorch~\citep{paszke2017automatic} and Scikit-learn~\citep{scikit-learn}.
This work was supported in part by the EPSRC Centre for Doctoral Training in Data Science, funded by the UK Engineering and Physical Sciences Research Council (grant EP/L016427/1) and the University of Edinburgh.
